---
title: "Text Mining in R"
author: "Teja Pristavec"
institute: "UVA BII Social & Decision Analytics Division"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "metropolis", "metropolis-fonts"]
    lib_dir: libs
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---
class: 

```{r, include = FALSE}
library(countdown)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)
```

# Let's get some text data

Follow along! Quick ways to get some text:
* Various books via [gutenbergr](https://cran.r-project.org/web/packages/gutenbergr/vignettes/intro.html)
* Song lyrics via [genius](https://github.com/josiahparry/genius)
* H.C. Andersen’s fairy tales via [hcandersenr](https://github.com/EmilHvitfeldt/hcandersenr)
* Proust via [proustr](https://colinfay.me/proustr-package/)
* Zen koans via [koanr](https://github.com/malcolmbarrett/koanr)
* Harry Potter via [harrypotter](https://github.com/bradleyboehmke/harrypotter)

`r countdown(minutes = 5, seconds = 00, play_sound = TRUE, box_shadow = FALSE, color_background = "#FAFAFA")`

---
class: 

# Let's get some text data

I'll be working with Kafka's Metamorphosis. 
<br>

```{r eval = TRUE, tidy = TRUE}
library(dplyr)
library(gutenbergr)

# gutenberg_metadata %>% filter(title == "Metamorphosis")
metamorphosis <- gutenberg_download(5200)
```
--
```{r eval = TRUE, tidy = TRUE}
metamorphosis$gutenberg_id <- NULL
tail(metamorphosis, 8)
```

---
class: center, middle, inverse

# First, some clean-up

---
# Tokenization

**Tokenization** parses the text into elements (or tokens) of interest. In this case, we want words as units of analysis.

```{r, tidy = TRUE}
library(tidytext)
metwords <- metamorphosis %>% unnest_tokens(word, text)
```
--
```{r, tidy = TRUE}
tail(metwords, 10)
```

---
# Removing stopwords
**Stopwords** are commonly used words (like _if_, _when_, _the_, _after_) that we typically (but not always!) want to ignore.
```{r}
data(stop_words)
metwords <- metwords %>% anti_join(stop_words, by = "word")
```
--
```{r, tidy = TRUE}
tail(metwords, 10)
```
---

class: center, middle, inverse

# Word frequencies

---
# Most frequent words

```{r, tidy = TRUE}
library(ggplot2, verbose = FALSE)

metwords %>% count(word, sort = TRUE) 

```

---
# Most frequent words

.pull-left[
```{r plot1, eval = FALSE}
library(ggplot2)

metwords %>%
  count(word, sort = TRUE) %>%
  filter(n > 30) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
     geom_col() +
     coord_flip() +
     labs(title = 
       "Most frequent words (n > 30)", 
       x = "Word", y = "Frequency (n)")
```
]
--
.pull-right[
```{r plot1-out, ref.label = "plot1", echo = FALSE}
```
]

---

class: center, middle, inverse

# Sentiments

---
# Sentiments

Tidytext comes with four sentiment lexicons: nrc, bing, afinn, loughran. Except for afinn, the lexicons have positive and negative sentiment categories. NRC also has anger, anticipation, disgust, joy, sadness, surprise, and trust; loughran also has "litigious", "uncertainty", "constraining", and "superfluous."

```{r}
bing <- get_sentiments("bing")
bing$sentiment <- as.factor(bing$sentiment)

metsent <- metwords %>%
  inner_join(bing, by = "word") %>%
  count(word, sentiment, sort = TRUE)  %>%
  ungroup()
```

---
# Sentiments

.pull-left[
```{r plot2, eval = FALSE}
library(wesanderson)

metsent %>%
 filter(n > 3) %>%
 mutate(word = reorder(word, n)) %>%
 ggplot(aes(word, n, fill = sentiment)) +
  geom_col() +
  coord_flip() + 
  facet_wrap(~ sentiment,
             scales = "free_y") +
  labs(title = 
   "Word frequency by sentiment (n > 3)",
    x = "Frequency", fill = "Sentiment") +
  scale_fill_manual(values = 
    wes_palette(n = 2, 
                name = "Moonrise3", 
                type = "discrete")) + 
  theme(legend.position = "bottom")
```
]
--
.pull-right[
```{r plot2-out, ref.label = "plot2", echo = FALSE}
```
]

---
class: center
# Sentiment over time

```{r,  include = FALSE}
library(tidyr)
cols <- c("#7FC97F", "#BEAED4", "#BEAED4")

# Bing
mettime_bing <- metwords %>%
  mutate(rowpos = row_number(), method = "Bing") %>%
  inner_join(bing, by = "word") %>%
  count(word, index = rowpos %/% 100, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative,
          sentcolor = ifelse(sentiment > 0, "positive",
                       ifelse(sentiment < 0, "negative",
                        ifelse(sentiment == 0, "neutral", NA))))

plotbing <- ggplot(mettime_bing, aes(index, sentiment, fill = sentcolor)) +
                  geom_col(show.legend = FALSE) +
                  scale_fill_manual(values = cols) +
                  labs(title = "Sentiment score per 100 words over time", x = "", y = "Bing")

# Nrc
mettime_nrc <- metwords %>%
  mutate(rowpos = row_number(), method = "NRC") %>%
  inner_join(filter(get_sentiments("nrc"), sentiment == "positive" | sentiment == "negative"), by = "word") %>%
  count(word, index = rowpos %/% 100, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative,
         sentcolor = ifelse(sentiment > 0, "positive",
                      ifelse(sentiment < 0, "negative",
                       ifelse(sentiment == 0, "neutral", NA))))

plotnrc <- ggplot(mettime_nrc, aes(index, sentiment, fill = sentcolor)) +
                  geom_col(show.legend = FALSE) +
                  scale_fill_manual(values = cols)  +
                  labs(x = "Word group", y = "NRC")
```

```{r}
library(cowplot); plot_grid(plotbing, plotnrc, ncol = 1, align = "v", axis = "l")
```

```{r}
plot_grid(plotbing, plotnrc, ncol = 1, align = "v", axis = "l")
```

---

class: center, middle, inverse

# Word clouds

WARNING! We don't like word clouds. See [this note](http://drewconway.com/zia/2013/3/26/building-a-better-word-cloud) about building wordclouds 2.0.

---
# Word clouds

.pull-left[
```{r plot3, eval = FALSE}
library(wordcloud)

metwords %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100, 
                 random.order = FALSE))
```
]

--
.pull-right[
```{r plot3-out, ref.label = "plot3", echo = FALSE}
```
]

---
# Word clouds

.pull-left[
```{r plot4, eval = FALSE}
library(reshape2)

metsent %>%
  acast(word ~ sentiment, value.var = "n", 
        fill = 0) %>%
  comparison.cloud(colors = c("darkblue", "brown"),
                   max.words = 100)
```
]

--
.pull-right[
```{r plot4-out, ref.label = "plot4", echo = FALSE}
```
]

---

class: center, middle, inverse

# N-grams

---
# N-grams

.pull-left[
```{r}
# Bi-grams
metbi <- metwords %>%
 unnest_tokens(bigram, word, token = "ngrams", n = 2)
metbi %>% count(bigram, sort = TRUE)
```
]

.pull-right[
```{r}
# Tri-grams
mettri <- metwords %>%
 unnest_tokens(trigram, word, token = "ngrams", n = 3)
mettri %>% count(trigram, sort = TRUE)
```
]

---

class: center, middle, inverse

# Word networks

---
# Word networks

.pull-left[
```{r plot5, eval = FALSE}
library(igraph); library(ggraph); library(widyr)

bigrams_separated <- metbi %>%
  separate(bigram, c("word1", "word2"), sep = " ")
bigram_counts <- bigrams_separated %>% 
  count(word1, word2, sort = TRUE) %>%
  filter(!is.na(word1) & !is.na(word2))
bigram_graph <- bigram_counts %>%
  filter(n > 2) %>%
  graph_from_data_frame()

bigram_plot <- grid::arrow(type = "closed", 
                     length = unit(.1, "inches"))

ggraph(bigram_graph, layout = "fr") +
 geom_edge_link(aes(edge_alpha = n), 
                show.legend = FALSE,
                arrow = bigram_plot, 
                end_cap = circle(.07, "inches")) +
 geom_node_point(color = "burlywood3", size = 3) +
 geom_node_text(aes(label = name), vjust = 1, 
                hjust = 1, repel = TRUE) +
 labs(title = "Metamorphosis bigrams") +
 theme_void()
```
]

--

.pull-right[
```{r plot5-out, ref.label = "plot5", echo = FALSE}
```
]

---

class: center, middle, inverse

# Word correlations

---
# Word correlations
.pull-left[
```{r}
# Create "units"
metwords <- metwords %>% mutate(section = row_number() %/% 50) %>%
                         filter(section > 0)

# Co-occuring words
metpairs <- metwords %>%
  pairwise_count(word, section, sort = TRUE)
metpairs
```
]

--

.pull-right[
```{r}
# Correlations
# Phi: H0 it is equally likely for either word to appear alone as it is that both of them or neither of them appear

metcors <- metwords %>%
           group_by(word) %>%
           filter(n() >= 20) %>%
           pairwise_cor(word, section, sort = TRUE)
metcors
```
]

---
# Word correlations

.pull-left[
```{r plot6, eval = FALSE}
# Words associated with...
metcors %>%
  filter(item1 %in% c("clerk", "door")) %>%
  group_by(item1) %>%
  top_n(5) %>%
  ungroup() %>%
  mutate(item2 = reorder(item2, correlation)) %>%
  ggplot(aes(item2, correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip()
```
]

--

.pull-right[
```{r plot6-out, ref.label = "plot6", echo = FALSE}
```
]

---
# Correlation graph

.pull-left[
```{r plot7, eval = FALSE}
metcors %>%
  filter(correlation > .15) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_void()
```
]

--

.pull-right[
```{r plot7-out, ref.label = "plot7", echo = FALSE}
```
]

---
class: center, middle, inverse

# Phew :)